---
title: "Collected Homework 3"
author: "Jackie Budka, Kristina Franklin, Connor Hathaway"
date: "10/25/2021"
output: html_notebook
---

## Questions
7.6, a-b
7.14
7.20
7.21
7.22


### Question 7.6, a-b
The data set storms is included in the dplyr package. It contains information about 425 tropical storms in the Atlantic.

a. Produce a histogram of the wind speeds in this data set. Fill your bars using the category variable so you can see the bands of color corresponding to the different storm categories.

b. Repeat part (a) but make a histogram of the pressure variable. You should observe that high category storms have low pressure.
```{r }

```

### Question 7.14
Make a scatterplot showing  
C
O
2
  uptake as a function of concentration level for the built-in data set CO2. Include a smoothed fit line and color by Type. Facet your plot to one plot for each Plant.
```{r }

```

### Question 7.20
Use the People data set from the Lahman package.

Create a barplot of the birth months of all players. (Hint: geom_bar())
Create a barplot of the birth months of the players born in the USA after 1970.
```{r }

```

### Question 7.21
Make a line graph of the total number of babies of each sex versus year.
```{r }

```

### Question 7.22
Make a line graph of the number of different names used for each sex versus year.
```{r }

```




