---
title: "Search and Rescues in the Adirondacks"
author: "Kristina Franklin, Rosie Delwiche, Connor Hathaway, Jackie Budka"
output:
  html_notebook:
    df_print: paged
    number_sections: yes
---

# Introduction

Every year, the New York State Forest Rangers have to rescue people who use the outdoors for recreation. Some get injured and need evacuation, some get lost and need search and rescue, but all put a burden on our park services resources. Any efforts to educate people on how to be safer and more responsible in nature will go a long way towards helping alleviate this burden but the Department of Environmental Conservation does not have the resources to market to everyone. In our analysis, we will try to identify groups that are at a greater risk of needing evacuation so we can make a recommendation on where best to allocate resources on awareness. We decided to focus on the Adirondack Park because of the region's high traffic and ability to attract inexperienced visitors. &&&&The variables of interest are the amount of rangers involved, amount of people being rescued, age and gender of rescued, and the type of activity that caused the accident. We will be analyzing the rescues happening in the Adirondack Park to try to find groups of people who are at a greater risk of needing rescue and would therefore benefit more from targeted awareness campaigns.

...

# Background

This is observational data originally found on Data World (https://data.world/) from the NYSDEC on forest ranger incident reports. In order to help understand the data it would be helpful for the reader to have previous knowledge about recreational activities in New York State forests and the risks involved with those activities. 

```{r message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(janitor)
library(lubridate)
library(tidymodels)
library(httr)
library(jsonlite)
library(sf)
library(tmap)
library (readr)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
urlfile="https://raw.githubusercontent.com/JaBudka/STAT383_F21/Project/update.csv"
raw_sr_data<-read.csv(url(urlfile)) %>%
  clean_names()
raw_adk_data <- raw_sr_data %>%
  filter(incident_adirondack_park == "TRUE")
```
...

# Methods and Results

To help us visualize the location found data we can look at all the incidents plotted on a map of New York State as shown in section 3.1. From visual inspection, we can see the highest density of rescues occur in the Adirondacks. We can verify this by using the table function to summarize the results.

```{r echo=FALSE}
count_adk <-  table(raw_sr_data['incident_adirondack_park'])
  rownames(count_adk) = c("Outside ADK", "Inside ADK")
count_adk
```

## Location Found of all Incidents
```{r  echo=FALSE, message=FALSE, warning=FALSE}
raw_sr_map <- raw_sr_data[complete.cases(raw_sr_data), ] %>%
st_as_sf(coords = c("location_found_longitude", "location_found_latitude"), crs = 4326)
tmap_mode("view")
tm_shape(raw_sr_map) +
  tm_dots(size=0.02,col="red", alpha = 0.5) + tm_legend(outside = TRUE) 
```

...

One of the variables we focused the most on was the age of the rescued. Section 3.2 displays the location found data again by subject age except this time just for incidents that started in the Adirondack park. There are a few outliers where people were found outside the park. These are likely due to the person being reported missing but then being found at home. 

## Location Found in Adirondacks Grouped by Age
```{r  echo=FALSE, message=FALSE, warning=FALSE}
adk_geom_data <- raw_adk_data[complete.cases(raw_adk_data), ] %>%
st_as_sf(coords = c("location_found_longitude", "location_found_latitude"), crs = 4326) 
tmap_mode("view")
tm_shape(adk_geom_data) +
  tm_dots(size=0.02,col="subject_age", alpha = 0.7, palette = "Spectral")
```

The data frame can be overwhelming to look at. It is easier to digest when summarized with the "table" function. We can use this function to better visualize the variables of interest. 

Some initial observations are that more men need assistance than women, there were more searches than rescues, recoveries, or fugitive searches combined, and the most common activity to need assistance was hiking, followed not so closely by boating.

```{r echo=FALSE}
count_gender <-  table(raw_adk_data['subject_gender'])
count_gender
count_rtype <- table(raw_adk_data['response_type'])
count_rtype
count_activity <- table(raw_adk_data['activity'])
count_activity

```

Our initial endeavor was to see if there was a correlation between subject age and type of response. From our analysis we concluded that there was no such correlation. The box plot is a helpful visualization because 

There seems to be a relationship between the subject's age and what type of response.  It can be concluded that as people get older, they may become more familiar with the land, or simply be more careful with their activities.  Search and Rescue responses are the only type that occur for people 30 and under, proving that the younger people should probably have more training on certain skills before traveling into the mountains alone.  Although, the mean is around 35 to 40 years old, meaning that mostly people over 30 are more common in general in the area, and therefore needing the help just as much.  Overall all people traversing into the mountains should have better safety awareness before going out alone, in case any problems occur.

Another important point to make about this data is the noticeable relation between older people and recovery.  As we all know, as we age our bodies are not as capable as they used to be, meaning they are more likely to be injured, causing a need to be rescued.  One way to decrease the need for rescues could be extra training about safety precautions and give fair warnings about certain activities.  For example if a hike has one area that gets slippery before the rest, put up more signs or make sure it is mentioned before anyone even begins the excursion.
```{r echo=FALSE}
raw_adk_data %>% 
  ggplot(aes(y = subject_age, x = response_type)) +
  geom_boxplot()+
  ggtitle("Subject Age vs Response Type") 
```


```{r echo=FALSE}
search_data <- raw_adk_data %>%
  filter(response_type=="Search")
rescue_data <- raw_adk_data %>%
  filter(response_type=="Rescue")
recovery_data <- raw_adk_data %>%
  filter(response_type=="Recovery")
MArecovery <- mean(recovery_data$subject_age, na.rm = "TRUE")
MArescue <- mean(rescue_data$subject_age, na.rm = "TRUE")
MAsearch <- mean(search_data$subject_age, na.rm = "TRUE")
cat('Mean ages
Recovery= ',MArecovery)
cat('
Rescue= ',MArescue)
cat('
Search= ',MAsearch)
```

```{r Load Data, include=FALSE}
raw_update <- read.csv("update.csv") %>%
  clean_names()
```

Just ADK
```{r include=FALSE}
xs <- raw_update %>%
  filter(incident_adirondack_park == "TRUE")
```


```{r include=FALSE}
y <- xs %>%
  filter(incident_time_elapsed>0)
```

Another variable that was considered was the time range to close the incident case.  The mean time per incident type was calculated, giving an average of how long each one took to complete.  Surprisingly, rescue took more time overall than search or recovery.  This is most likely because the case itself is longer in general. For search, when they find the person, the case is closed.  But, when a rescue is required, they need more time to relieve the person from the situation, and get them to safety.
A way to decrease this time could be to change how they decide search, rescue and recovery.  If search was simply finding the person, rescue is bringing them out, and recovery is their time in the hospital, then many more of these situations can be applied to each incident.

```{r echo=FALSE}
mean(y$incident_time_elapsed)

search <- y %>%
  filter(response_type=="Search")
rescue <- y %>%
  filter(response_type=="Rescue")
recovery <- y %>%
  filter(response_type=="Recovery")

rev <- mean(search$incident_time_elapsed, na.rm = "TRUE")
res <- mean(rescue$incident_time_elapsed, na.rm = "TRUE")
sea <- mean(recovery$incident_time_elapsed, na.rm = "TRUE")
cat('Mean incident time elapsed
Recovery= ',rev)
cat('
Rescue= ',res)
cat('
Search= ',sea)
```


Perform at least one relevant hypothesis test. 



The first hypothesis test was a two-tailed test to find the difference between between amount of males and females.

```{r echo=FALSE}
female <- raw_adk_data %>%
  filter(subject_gender == "F")

male <- raw_adk_data %>%
  filter(subject_gender == "M")

h1 <- t.test(female$subject_age, male$subject_age, alternative = "two.sided", var.equal = FALSE)
h1
```

The second hypothesis test performed was a single-tailed hypothesis to see if the ages between rescued males and females differ.  
The null hypothesis is mu_f - mu_m = 0
The alternative hypothesis is mu_f - mu_m < 0
The t-test is performed to find the difference between the two samples.
After the t-test is run, the value is -3.176, meaning we reject the null hypothesis because the difference between males and females is not 0.

```{r echo=FALSE}

female <- raw_adk_data %>%
  filter(subject_gender == "F")

male <- raw_adk_data %>%
  filter(subject_gender == "M")

h2 <- t.test(female$subject_age, male$subject_age, alternative = "less", var.equal = FALSE)
h2

```


Since two variable relations were considered, another set of t-tests were performed.  A two sided t-test where mu = 
```{r echo=FALSE}
# Does the mean case time differ between search and rescue?
t.test(search$incident_time_elapsed,rescue$incident_time_elapsed,alternative = "two.sided",conf.level = .98)

# Does the mean case time differ between recovery and search?
t.test(recovery$incident_time_elapsed,search$incident_time_elapsed,alternative = "two.sided",conf.level = .98)

```
```{r echo=FALSE}
incident_model <- lm(incident_time_elapsed~number_of_rangers_involved, data = y)
incident_model
# intercept 1206.5
# slope 624.2 
# this means predicted time = 624.2 * rangers involved
```

```{r echo=FALSE}

#y %>% ggplot(aes(x = number_of_rangers_involved, y = incident_time_elapsed)) +
#  geom_point() +
#  geom_abline(intercept = 3.257e+00, slope = 8.195e-06 )
#incident_model$residuals
sum(incident_model$residuals^2)
summary(incident_model)

# Because p is less than alpha, we reject the null hypothesis. We have reason to believe that there is a linear relationship between incident time elapsed and number of rangers involved
```



Check the various assumptions of for statistical tests.

The following outputs are from the linear model made for number of rangers involved based on the subject age. In this test, we wanted to see if we could correlate the amount of rangers that a case required to the age of the subject. From the Q-Q plot, we can conclude that the data is not normal because the standardized residuals do not follow the dotted linear line. 

```{r echo=FALSE}
model = lm(number_of_rangers_involved ~ subject_age, data = raw_adk_data)
summary(model)
plot(model)
```

```{r}

# predict the time to close a case with 3 rangers
predict(incident_model, newdata = data.frame(number_of_rangers_involved = 3))

# correlation between time elapsed and number of rangers for all types of incidents
y %>%
  ggplot(aes(x = incident_time_elapsed, y = number_of_rangers_involved, color = response_type)) +
  geom_point(size = 0.1) +
  facet_wrap(vars(response_type))

# looking at correlation for each response type
# there is a high correlation between time elapsed and number of rangers involved for fugitive search
# the other ones dont show a high correlation but this is kinda expected because there are lots of outliers
y %>%
  group_by(response_type) %>%
  summarize(r = cor(incident_time_elapsed, number_of_rangers_involved, use = "complete.obs"))


# incident model qq plot
plot(incident_model)

```





For the linear regression analysis, interpret coefficients and/or make relevant predictions and
summarize their meaning.

Below we create a plot of number of rangers involved based on the subject age. Remember that as previously stated, the data is not normal based on the Q-Q plot and therefore does not have statistical significance, however we were able to complete an analysis of the linear regression line. 

```{r echo=FALSE}
raw_adk_data %>% 
  ggplot(aes(x = subject_age, y = number_of_rangers_involved))+
  geom_point()+
  geom_abline(intercept = 3.142535, slope = 0.004627, col="magenta")+
  ggtitle("Rangers to Age Regression") 
```


The following output is the linear model summary. The intercept for this linear regression is 3.142535 and
The β1 value or slope of this linear regression is 0.004627 which also is very close to zero under the scope of this data. To check correlation of the data, we perform a hypothesis test with the null hypothesis as β1=0 and the alternative hypothesis is β1≠0, with a level of significance of alpha = 0.05, we find our P-value is 0.375 which is greater than our level of significance of 0.05 and therefore we fail to reject the null hypothesis. This means that Beta1 is equal to zero because the value of 0.004627 is not statistically significant which means that there is not a correlation between subject age and number of rangers involved. 


```{r echo=FALSE}
x <- lm(formula = number_of_rangers_involved ~ subject_age,data=raw_adk_data)
summary(x)
```

We also calculated the correlation of this data below using the "cor()" function. 
```{r}
cor(raw_adk_data$subject_age,raw_adk_data$number_of_rangers_involved, use = "complete.obs")
```

Below we compute the correlation of number of rangers involved based on subject age, but in this case we group by response type. 

```{r}

raw_adk_data %>%
  group_by(response_type) %>%
  summarize(r = cor(x = subject_age, y = number_of_rangers_involved, use = "complete.obs"))
```



```{r echo=FALSE}

#y %>% ggplot(aes(x = number_of_rangers_involved, y = incident_time_elapsed)) +
#  geom_point() +
#  geom_abline(intercept = 3.257e+00, slope = 8.195e-06 )
#incident_model$residuals
sum(incident_model$residuals^2)
summary(incident_model)

# Because p is less than alpha, we reject the null hypothesis. We have reason to believe that there is a linear relationship between incident time elapsed and number of rangers involved
```
...


# Conclusions

We were unable to find a demographic to target with an awareness campaign because of low statistical significance. From a common sense standpoint it would make sense to market more to hikers because they make up the vast majority of those in need of assistance but since we do not have data on how many people participate in each activity, we cannot definitively say that hikers need assistance at a higher rate than those in other activities. We could extend this study to the rest of New York State and even the rest of the United States, but not all regions are comparable due to differences in usage and common activities. What works in one area might not work elsewhere. We also noticed some errors in the data, specifically in the time elapsed. This was most likely due to user error. The DEC could be notified that their method of inputting dates is flawed.
One thing that would improve this study is more data and general usage data to compare rates of needed rescue in different demographics. The data on rescues during Covid would be especially interesting to look at since more people engaged with the outdoors. Specifically more inexperienced people. It would also be informative to have another variable in the study that qualitatively rates victim's level of competence. It would be expected that almost all cases are due to inexperienced people but it could also be possible that highly competent people did not realize their limit.

...


# References {-}

Data.world
https://data.world/data-ny-gov/u6hu-h7p5
